Actuator line modeling of vertical-axis turbines

### 2016-04-29 paper.tex
Used --> modeled
Add unsteadiness to theory section; resolves #10

    \caption{Drawings of the turbines used in this study: the UNH-RVAT (left)
    and DOE/SNL RM2 (right).}

@@

The blade lift and drag forces are calculated as
\begin{equation}
    F_l = \frac{1}{2} \rho A_\mathrm{elem} C_l |\vec{U}_\mathrm{rel}|^2,
\end{equation}
and
\begin{equation}
    F_d = \frac{1}{2} \rho A_\mathrm{elem} C_d |\vec{U}_\mathrm{rel}|^2,
\end{equation}
where $\rho$ is the fluid density, $A_\mathrm{elem}$ is the blade element
planform area (span $\times$ chord), $\vec{U}_\mathrm{rel}$ is the local
relative velocity, and $C_l$ and $C_d$ are the sectional lift and drag
coefficients, linearly interpolated from a table per the local angle of attack.
The forces are then projected onto the rotor coordinate system to calculate
torque, overall drag, etc. Forces from the turbine shaft and blade support
struts are computed in a similar way. After the force on the actuator lines from
the flow is computed, it is then added to the Navier--Stokes equations as a body
force or momentum source (per unit density, assuming incompressible flow):
\begin{equation}
    \frac{\mathrm{D} \vec{u}}{\mathrm{D} t} = - \frac{1}{\rho} \nabla p + \nu
    \nabla^2 \vec{u} + F_\mathrm{turbine}.
\end{equation}

>>>>

    \caption{Drawings of the turbines modeled in this study: the UNH-RVAT (left)
    and DOE/SNL RM2 (right).}

@@

Inspecting Figure~\ref{fig:geom-alpha-urel}, it is apparent that CFTs are
inherently unsteady machines, encountering large amplitude and rapid
oscillations in angle of attack, often exceeding the rotor blades' static stall
angles in typical operation. The unsteadiness can be characterized by a reduced
frequency~\cite{Leishman2006}
\begin{equation}
    k = \frac{\omega c}{2 U_\infty},
\end{equation}
which assumes the free stream velocity is constant. Unsteady effects begin to
become significant for $k > 0.05$, and can become dominant for $k \ge 0.2$.

For a VAT or CFT reduced frequency can be reformulated in terms of the tip speed
ratio as
\begin{equation}
    k = \frac{\lambda c}{2R},
\end{equation}
which is then also a function of solidity or chord-to-radius ratio $c/R$. As an
example, a large scale, relatively low solidity Darrieus wind turbine such as
the Sandia 34 m diameter Test Bed, with an equatorial blade chord of 0.91 m
\cite{Murray2011}, a reduced frequency $k=0.16$ is encountered based solely on
angle of attack oscillations at $\lambda=6$. For a smaller scale CFT, e.g., with
$c/R = 0.25$, operating at $\lambda = 2$, the reduced frequency is 0.25. It
follows that unsteady effects will be significant for cross-flow turbines even
in the absence of nonlinear effects such as stall.

Assuming unsteady effects can be appropriately modeled, the blade element lift
and drag forces are calculated as
\begin{equation}
    F_l = \frac{1}{2} \rho A_\mathrm{elem} C_l |\vec{U}_\mathrm{rel}|^2,
\end{equation}
and
\begin{equation}
    F_d = \frac{1}{2} \rho A_\mathrm{elem} C_d |\vec{U}_\mathrm{rel}|^2,
\end{equation}
where $\rho$ is the fluid density, $A_\mathrm{elem}$ is the blade element
planform area (span $\times$ chord), $\vec{U}_\mathrm{rel}$ is the local
relative velocity, and $C_l$ and $C_d$ are the sectional lift and drag
coefficients, linearly interpolated from a table per the local angle of attack.
The forces are then projected onto the rotor coordinate system to calculate
torque, overall drag, etc. Forces from the turbine shaft and blade support
struts are computed in a similar way. After the force on the actuator lines from
the flow is computed, it is then added to the Navier--Stokes equations as a body
force or momentum source (per unit density, assuming incompressible flow):
\begin{equation}
    \frac{\mathrm{D} \vec{u}}{\mathrm{D} t} = - \frac{1}{\rho} \nabla p + \nu
    \nabla^2 \vec{u} + F_\mathrm{turbine}.
\end{equation}
